---
title: "EDA"
author: "Crystal Chen & Sara Golestaneh"
output: pdf_document
---

The following is the exploratory data analysis completed for Project 2 ("Predicting Re-admission for Diabetic Patients"). To see R code, please view the Rmarkdown file.

```{r, include=FALSE}
library(tidyverse)
library(here)
# reading in data
d = read.csv(here("combined.csv"))
```

## Race vs. Re-admission: 

```{r, echo = F}
# looking at race in relation to re-admission
levels(d$race) = c("Other", "AfricanAmerican", "Asian", "Caucasian",
                   "Hispanic", "Other")
racedf = d %>% group_by(race) %>% summarize(n = n())
```

```{r, echo = F, fig.height=3}
# piechart of proportion of various races
ggplot(data = racedf, aes(x = "", y = n, fill = race)) + 
  geom_bar(stat = "identity", width = 1) + 
  coord_polar("y", start = 0) + 
  scale_fill_manual(values = c("lightblue4", "lightpink4", 
                               "lightyellow2", "lavenderblush3", "lightseagreen")) + 
  ggtitle("    Distribution of Races") + 
  labs(fill = "Race") + theme_void()
```

Figure 1: A piechart displaying the distribution of patient's race in entire dataset
$\newline$

```{r, echo = F, fig.height=4, fig.width=4}
# proportion of readmission for each race
racedf = d %>% group_by(race, readmitted_.30) %>% summarize(n = n()) %>% 
  group_by(race) %>% mutate(prop = n/sum(n))
racedf$readmitted_.30 = c(rbind(rep("No", 5), rep("Yes", 5)))

ggplot(data = racedf, aes(x = race, y = prop, fill = readmitted_.30)) + 
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("lightblue2", "lightpink4")) + 
  ggtitle("Proportion Re-admission by Race") + 
  theme(plot.title = element_text(size = 12), 
        axis.text.x = element_text(size = 12, angle = 90, hjust = 1, vjust = 0.5), 
        legend.text = element_text(size = 12), 
        legend.title = element_text(size = 12)) + 
  labs(fill = "Re-admission", x = "Race", y = "Counts")
```

Figure 2: Proportion of readmitted vs. non-readmitted patients in each group

## Gender vs. Re-admission

```{r, echo = F}
# looking at gender vs. re-admission
genderdf = d %>% group_by(gender, readmitted_.30) %>% summarize(n = n())
genderdf$readmitted_.30 = c("No", "Yes", "No", "Yes")
```

```{r, echo = F, fig.height=3, fig.width=4}
# frequency of re-admission for each gender type
ggplot(data = genderdf, aes(x = gender, y = n, fill = readmitted_.30)) + 
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("lightblue2", "lightpink4")) + 
  ggtitle("Frequency of Re-admission by Gender") + 
  theme(plot.title = element_text(size = 12), axis.text = element_text(size = 12), 
        axis.title = element_text(size = 12), legend.text = element_text(size = 12), 
        legend.title = element_text(size = 12)) + 
  labs(fill = "Re-admission", x = "Gender", y = "Counts")
```

Figure 3: Frequency of re-admission by gender

## Age vs. re-admission

```{r, echo = F}
# distribution of age vs. re_admission
agedf = d %>% group_by(age, readmitted_.30) %>% summarize(n = n()) %>% 
  group_by(age) %>% mutate(prop = n/sum(n))
agedf$age = as.factor(agedf$age)
agedf$readmitted_.30 = c(rbind(rep("No", 10), rep("Yes", 10)))
```

```{r, echo = F, fig.height=4, fig.width=4}
ggplot(data = agedf, aes(x = age, y = n)) + geom_bar(stat = "identity") +
  scale_fill_manual(values = c("lightblue4")) + 
  ggtitle("Distribution of Patients' Age") + 
  theme(plot.title = element_text(size = 12), 
        axis.text.x = element_text(size = 12)) + 
  labs(fill = "Re-admission", x = "Age", y = "Counts")
```

Figure 4: Distribution of patients' age
$\newline$

```{r, echo = F, fig.height=3, fig.width=4}
ggplot(data = agedf, aes(x = age, y = prop, fill = readmitted_.30)) + 
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("lightblue2", "lightpink4")) + 
  ggtitle("Proportion of Re-admission by Age") + 
  theme(plot.title = element_text(size = 12), 
        axis.text.x = element_text(size = 12), 
        legend.text = element_text(size = 12), 
        legend.title = element_text(size = 12)) + 
  labs(fill = "Re-admission", x = "Age", y = "Proportion")
```

Figure 5: Proportion of re-admission within each age group

## Medication and Procedures vs. Re-admission 

```{r, echo = F}
# distribution of various procedures against readmission  
proceduresdf = d %>% select(num_lab_procedures, num_procedures, readmitted_.30, 
                            num_medications) %>% 
  mutate(num_tot_procedures = num_lab_procedures + num_procedures)
proceduresdf$readmitted_.30 = factor(ifelse(proceduresdf$readmitted_.30 == 0, "No", "Yes"))
```

```{r, echo = F, fig.height=2}
ggplot(data = proceduresdf, aes(x = readmitted_.30, y = num_medications)) + 
  geom_boxplot(fill = c("lightblue1", "lightpink4")) + coord_flip() + 
  ggtitle("Boxplot of Number of Medications") + 
  theme(plot.title = element_text(size = 12), 
        axis.text.x = element_text(size = 12)) + 
  labs(x = "Readmission", y = "Number of Medications")
```

Figure 6: Boxplot of number of medications for each re-admission type
$\newline$

```{r, echo = F, fig.height=2}
ggplot(data = proceduresdf, aes(x = readmitted_.30, y = num_tot_procedures)) + 
  geom_boxplot(fill = c("lightblue1", "lightpink4")) + coord_flip() + 
  ggtitle("Boxplot of Total Number of Procedures") + 
  theme(plot.title = element_text(size = 12), 
        axis.text.x = element_text(size = 12)) + 
  labs(x = "Readmission", y = "Number of Procedures")
```

Figure 7: Boxplot of total number of procedures for each re-admission type

## Diagnosis 1, 2, 3, vs. Re-admission 

```{r, echo = F}
# function to generate appropriate bargraphs given proper dataset
diagnostic_graphs = function(df, diagnosis, title, xlab){
  ggplot(data = df, aes(x = diagnosis, y = prop, fill = Readmittance_Type)) +
    geom_bar(stat = "identity") + 
    scale_fill_manual(labels = c("Not readmitted", ">30 days", "<30 days"), 
                      values = c("lightblue1", "lavenderblush3", "lightpink4")) + 
    ggtitle(title) + 
    theme(plot.title = element_text(size = 12), 
          axis.text.x = element_text(size = 12, angle = 90, hjust = 1, vjust = 0.5), 
          legend.text = element_text(size = 12), 
          legend.title = element_text(size = 12)) + 
    labs(fill = "Readmittance Type", x = xlab, y = "Proportion")
}
```

```{r, echo = F, fig.height=2.5, fig.width=4}
diag1 = d %>% select(diag_1, readmitted_.30, readmitted_.30.1, readmitted_NO)
colnames(diag1) = c("diag_1", "RA_under_30", "RA_over_30", "RA_No")
diag1 = diag1 %>% 
  pivot_longer(RA_under_30:RA_No, names_to = "Readmittance_Type", values_to = "Value")
prop_diag1 = diag1 %>% 
  group_by(diag_1, Readmittance_Type) %>% 
  summarize(n = sum(Value)) %>% 
  group_by(diag_1) %>% 
  mutate(prop = n/sum(n))

diagnostic_graphs(df = prop_diag1, diagnosis = prop_diag1$diag_1,
                  title = "Readmittance Prop for Most Common Diag 1 Outcomes", 
                  xlab = "Diagnosis 1 Outcomes")
```

Figure 8: Stacked bargraph of proportion of re-admission for top 10 most frequent outcomes of diagnosis 1
$\newline$

```{r, echo = F, fig.height=2.5, fig.width=4}
# diagnostic 2
diag2 = d %>% select(diag_2, readmitted_.30, readmitted_.30.1, readmitted_NO)
colnames(diag2) = c("diag_2", "RA_under_30", "RA_over_30", "RA_No")
diag2 = diag2 %>% 
  pivot_longer(RA_under_30:RA_No, names_to = "Readmittance_Type", values_to = "Value")
prop_diag2 = diag2 %>% 
  group_by(diag_2, Readmittance_Type) %>% 
  summarize(n = sum(Value)) %>% 
  group_by(diag_2) %>% 
  mutate(prop = n/sum(n))

diagnostic_graphs(df = prop_diag2, diagnosis = prop_diag2$diag_2,
                  title = "Readmittance Prop for Most Common Diag 2 Outcomes", 
                  xlab = "Diagnosis 2 Outcomes")
```

Figure 9: Stacked bargraph of proportion of re-admission for top 10 most frequent outcomes of diagnosis 2
$\newline$

```{r, echo = F, fig.height=2.5, fig.width=4}
# diagnostic 3
diag3 = d %>% select(diag_3, readmitted_.30, readmitted_.30.1, readmitted_NO)
colnames(diag3) = c("diag_3", "RA_under_30", "RA_over_30", "RA_No")
diag3 = diag3 %>% 
  pivot_longer(RA_under_30:RA_No, names_to = "Readmittance_Type", values_to = "Value")
prop_diag3 = diag3 %>% 
  group_by(diag_3, Readmittance_Type) %>% 
  summarize(n = sum(Value)) %>% 
  group_by(diag_3) %>% 
  mutate(prop = n/sum(n))
diagnostic_graphs(df = prop_diag3, diagnosis = prop_diag3$diag_3,
                  title = "Readmittance Prop for Most Common Diag 3 Outcomes", 
                  xlab = "Diagnosis 3 Outcomes")
```

Figure 10: Stacked bargraph of proportion of re-admission for top 10 most frequent outcomes of diagnosis 3